Litrepl: Literate Paper Processor Promoting Transparency More Than Reproducibility

### 2025-01-18 paper/tex/paper.tex
Update papers

\begin{comment}
\begin{sh}
echo '\begin{verbatim}'
echo '$ cat input.tex | litrepl'
echo "sys.platform='linux'" | litrepl repl python >/dev/null
cat input.tex | litrepl
echo '\end{verbatim}'
\end{sh}
\end{comment}
%lresult
\begin{verbatim}
$ cat input.tex | litrepl
\begin{python}
import sys
print(f"I use {sys.platform} btw!")
\end{python}
\begin{result}
I use linux btw!
\end{result}
\end{verbatim}
%lnoresult

@@

\begin{figure*}[hbt!]
  \centering
  \begin{minipage}{\textwidth}
  %lresult
  \begin{verbatim}
  \begin{result}
  ... some output ...
  [LR:/tmp/litrepl/python/partial_7b81e1e.txt]
  \end{result}
  \end{verbatim}
  %lnoresult
  \end{minipage}
  \caption{Example partial result section that ends with a continuation tag,
  where \texttt{LR:} is a distinguishable prefix, and the rest is the filename
  storing a response of the interpreter to the section being evaluated.}
  \label{fig:partial-result}
\end{figure*}

@@

\begin{comment}
% lai

Lets write the "Discussion" section. of the paper. Please check the grammar of
the following points and make a linked narrative out of them. Please use Latex
markup.

---

The technical decision to abstract interpreters using text streams has its
positive and negative sides. While simplicity is the obvious bonus, the negatvie
consequences can be groouped into (1) No parallel evaluation at the
communication level. Th interpreter gets locked until it finishes evaluation of
one snippet before it can proceed to another one. (2) Transferrable data type is
limited to text-only streams.

We argue that the lack of parallel execution at the communication level can be
compensated by using interpreter-specific parallelism where and if it is
supported. For example, Python programs can use all sort of subprocess
utilities while shell programs have full access to shell job control.

Restricting streams to text-only data type seems to be a more funcdamental
limitation. Litrepl lifts this limitation by supporting text-only document
formats. Both Latex and Markdown have means of including rich data but they
don't encode these data directly. Instead they rely on side channels: file
system or network for referring such objects. Litrepl shares both pros such as
good representation in well-known version control systems, and cons, e.g.
requirement to organize such a side-channel data transfer explicitly.

Another controversial technical solution - to transfer the whole document at
input and output - can have performance penalties. In our experience the system
works sufficienlty good for few-thousand line documents, but for larger
documents delays might become uncomfortable even on modern computers. Still we
prefer to keep this interface to make editor integration as simple as possible:
typical plugin can just pipe the whole document through the tool using few lines
of code.

In addition, we allow a more performant integrations which take responsibility
for pre-parsing and piping only the relevant parts of the document. For these
approach, Litrepl offers \verb|print-regexp| command printing the anchor regexp
in a few common formats.
% lnoai
\end{comment}

% lresult
The technical decision to abstract interpreters using text streams comes with
both advantages and disadvantages. A key advantage is simplicity. However, there
are notable negative aspects. First, there is no parallel evaluation at the
communication level, meaning the interpreter is locked until it completes the
evaluation of one snippet before proceeding to the next. Second, the
transferable data type is restricted to text-only streams.

@@

A more performance-oriented integrations can make pre-parsing and pipe only
relevant document parts. For these approaches, Litrepl offers the
\verb|print-regexp| command, which outputs the anchor regexp in several common
formats.
% lnoresult

>>>>

\begin{comment}
\begin{sh}
litrepl restart python
echo '\begin{verbatim}'
echo '$ cat input.tex | litrepl'
echo "sys.platform='linux'" | litrepl repl python >/dev/null
cat input.tex | litrepl
echo '\end{verbatim}'
\end{sh}
\end{comment}
%lresult
\begin{verbatim}
$ cat input.tex | litrepl
\begin{python}
import sys
print(f"I use {sys.platform} btw!")
\end{python}
\begin{result}
I use linux btw!
\end{result}
\end{verbatim}
%lnoresult

@@

\begin{figure*}[hbt!]
  \centering
  \begin{minipage}{\textwidth}
  %lresult
  \begin{verbatim}
  \begin{result}
  ... some output ...
  [LR:/tmp/litrepl/python/partial_c335adc.txt]
  \end{result}
  \end{verbatim}
  %lnoresult
  \end{minipage}
  \caption{Example partial result section that ends with a continuation tag,
  where \texttt{LR:} is a distinguishable prefix, and the rest is the filename
  storing a response of the interpreter to the section being evaluated.}
  \label{fig:partial-result}
\end{figure*}

@@

The technical decision to abstract interpreters using text streams comes with
both advantages and disadvantages. A key advantage is simplicity. However, there
are notable negative aspects. First, there is no parallel evaluation at the
communication level, meaning the interpreter is locked until it completes the
evaluation of one snippet before proceeding to the next. Second, the
transferable data type is restricted to text-only streams.

@@

A more performance-oriented integrations can make pre-parsing and pipe only
relevant document parts. For these approaches, Litrepl offers the
\verb|print-regexp| command, which outputs the anchor regexp in several common
formats.
